### 2018-09-15 ICRA19/covControl.tex
shortened abstract, made experiments correspond with rest of paper

\subsection{Efficient control of correlation}
\begin{figure}[htb]
\begin{center}
	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:ControlledCorrelation}
Analytical results comparing maximum correlation $\rho_{xy}$. Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $\ell$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
}\end{figure}

@@

\section{Experiments}\label{sec:Experiments}
This section reviews three demonstrations of efficient shape control of many particles.
\subsection{Hardware experiment: particle angle of repose}
In this section we varied the approach angle to shape a swarm with a characteristic angle of repose.  
We then showed how the angle of repose can be increased by decreasing the magnitude of the force that moves the particles.
%
\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{aorPlot.pdf}
\end{center}
\vspace{-1.5em}
\caption{\label{fig:angleOfReposeExp}
Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity.
}\vspace{-1.5em}
\end{figure}
%
%\begin{figure}
%\begin{center}
%	\includegraphics[width=\columnwidth]{ironAndSaltCloseup.pdf}
%\end{center}
%\caption{\label{fig:ironAndSaltCloseup}
%Magnified image of salt and iron at angle of repose.
%}
%\end{figure}
%
The particles used are iron filings (Dowling Magnets, \#731019, $\approx0.5$~mm diameter), and the motive force is gravity. 
The particles were poured into a frame composed of two sheets of acrylic with 4.5 mm spacing between. Five 40 mm rods were epoxied between the sheets at angles 
 $\{0^\circ,15^\circ, 30^\circ, 45^\circ, 60^\circ \}$, generating angle of attacks $\beta = \{90^\circ,75^\circ, 60^\circ, 45^\circ, 30^\circ \}$.
 The frame was tested at three tilt angles with respect to horizontal: $\beta = \{90^\circ,60^\circ, 45^\circ \}$, resulting in $\{100,87, 71\}\%$ of the gravitational force.
 For each test the bottom of the frame was removed to allow unsupported particles to escape and the shape of the remaining particles was measured from photographs taken by a tripod-mounted camera inclined at the same tilt angle as the frame.
  The decreasing tilt angles resulting in increasing angle of repose values: $\alpha = \{42^\circ,50^\circ, 63^\circ \}$
These experiments show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle.  These result in varying levels of force and torque on the rod according to \eqref{eq:ForceAOR} and \eqref{eq:TorqueAOR}.

\subsection{Hardware experiment: Kilobot angle of repose}
We varied the approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose. 
 A Kilobot is a three-centimeter diameter,  low-cost robot. In our experiments, these robots were programmed to go toward the brightest light in the room. The experimental setup is shown in Fig. \ref{fig:SetUp}, also showing angle of repose of the Kilobots with  the pink rod at three angles of attack: 90$^\circ$, 67.5$^\circ$ and 45$^\circ$. The green lines show the triangle shape the robots pile up into. The cyan star is the center of mass of the rod, and the mean position and variance of the robots are shown with  red stars and red ellipses.
\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{SetUp.pdf}
\end{center}
\vspace{-1.5em}
\caption{\label{fig:SetUp}
Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks.
}\vspace{-1.5em}
\end{figure}

@@

\subsection{Hardware experiment: control of covariance}
To demonstrate covariance control, up to 100 Kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.  The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm. The system has an approximately 10 s lag time before the covariance switches direction.

\begin{figure}[!htb]
\begin{center}
\vspace{-1em}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-2em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering $\approx 50$ Kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The measured covariance is shown for five trials (different line colors). Frames above the plot show output from machine vision system and overlaid goal and measured covariance ellipses.
\vspace{-1.5 em}
}
\end{figure}

>>>>

\subsection{Efficient control of correlation  $\rho_{xy}$}\label{subsec:efficientCorrelation}
\begin{figure}[htb]
\begin{center}
	\includegraphics[width=\columnwidth]{ControlledCorrelation.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:ControlledCorrelation}
Analytical results comparing maximum correlation $\rho_{xy}$. Lines labeled \textsf{F} use the boundary layer friction model of \eqref{eq:correlationFriction} with four boundary layer thicknesses $\ell$ and the stable triangular configuration \eqref{eq:GravityCorrelation}.
}\end{figure}

@@

\section{Experiments}\label{sec:Experiments}
This section reviews three demonstrations of efficient shape control of many particles.
\subsection{Hardware experiment: particle angle of repose}
In this section we varied the approach angle to shape a swarm with a characteristic angle of repose.  
We then showed how the angle of repose can be increased by decreasing the magnitude of the force that moves the particles.
%
\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{aorPlot.pdf}
\end{center}
\vspace{-1.5em}
\caption{\label{fig:angleOfReposeExp}
Experimental data using iron particles, varying angle of attack, and three tilt angles to vary the force applied by gravity.
}\vspace{-1.5em}
\end{figure}
%
%\begin{figure}
%\begin{center}
%	\includegraphics[width=\columnwidth]{ironAndSaltCloseup.pdf}
%\end{center}
%\caption{\label{fig:ironAndSaltCloseup}
%Magnified image of salt and iron at angle of repose.
%}
%\end{figure}
%
The particles used are iron filings (Dowling Magnets, \#731019, $\approx0.5$~mm diameter), and the motive force is gravity. 
The particles were poured into a frame composed of two sheets of acrylic with 4.5 mm spacing between. Five 40 mm rods were epoxied between the sheets at angles 
 $\{0^\circ,15^\circ, 30^\circ, 45^\circ, 60^\circ \}$, generating angle of attacks $\beta = \{90^\circ,75^\circ, 60^\circ, 45^\circ, 30^\circ \}$.
 The frame was tested at three tilt angles with respect to horizontal: $\beta = \{90^\circ,60^\circ, 45^\circ \}$, resulting in $\{100,87, 71\}\%$ of the gravitational force.
 For each test the bottom of the frame was removed to allow unsupported particles to escape and the shape of the remaining particles was measured from photographs taken by a tripod-mounted camera inclined at the same tilt angle as the frame.
  The decreasing tilt angles resulted in increasing angle of repose values: $\alpha = \{42^\circ,50^\circ, 63^\circ \}$, corresponding to Fig.~\ref{fig:AngleOfReposeForce}(a).
%%% We removed these because they seemed unnecessary
% These experiments show how the pile of particles can be used to convey two scalar values: the force applied and the attack angle. 
% These result in varying levels of force and torque on the rod according to \eqref{eq:ForceAOR} and \eqref{eq:TorqueAOR}.

\subsection{Hardware experiment: Kilobot angle of repose}
We varied the approach angle to shape a swarm of Kilobots \cite{rubenstein2014programmable} with a characteristic angle of repose. 
 A Kilobot is a 3 cm diameter,  low-cost robot. In our experiments, these robots were programmed to go toward the brightest light in the room. The experimental setup is shown in Fig. \ref{fig:SetUp}, also showing angle of repose of the Kilobots with  the pink rod at three angles of attack: 90$^\circ$, 67.5$^\circ$ and 45$^\circ$, generating three different shapes, as in Fig.~\ref{fig:angle}. The green lines show the triangle shape the robots pile up into. The cyan star is the center of mass of the rod, and the mean position and variance of the robots are shown with  red stars and red ellipses.
\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{SetUp.pdf}
\end{center}
\vspace{-1.5em}
\caption{\label{fig:SetUp}
Steering Kilobots with lights and characteristic shape formed by Kilobot angle of repose for three angle of attacks.
}\vspace{-1.5em}
\end{figure}

@@

\subsection{Hardware experiment: control of correlation}
To demonstrate the correlation control of Sec.~\ref{subsec:efficientCorrelation}, 34 Kilobots were placed on a workspace with boundary layer thickness $\ell \approx 2$ cm and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.  The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm. The system has an approximately 10 s lag time before the covariance switches direction.

\begin{figure}[!htb]
\begin{center}
\vspace{-1em}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-2em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering $34$ Kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The measured covariance is shown for five trials (different line colors). Frames above the plot show output from machine vision system and overlaid goal and measured covariance ellipses.
\vspace{-1.5 em}
}
\end{figure}

### 2018-09-15 ICRA19/theory.tex
shortened abstract, made experiments correspond with rest of paper

Consider a swarm of granular particles applying force to a rod. 
If the rod moves slower than the particles, some particles will slide past the rod, but other particles will build up behind the rod in a characteristic triangular shape %defined by an apex angle. 
 defined by a steepest angle of descent perpendicular to the direction of particle motion. 
This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack $\alpha$ is the \emph{angle of repose}\footnote{angle of repose is a 3D phenomenon, but this paper only considers the 2D projection of the piled particles.}. 
 Three different values of angle of repose is shown in Fig.~\ref{fig:angle}. The center of mass of the rod is in the middle of the rod, but center of mass of the granular particles changes for different values of angle of repose. % for the media, see \cite{angleofRepose}. 
 By measuring the angle of repose for the particles shown in the top plot of Fig.~\ref{fig:AngleOfReposeForce}, we can estimate the force and torque that the swarm is applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
 In this plot, particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$ axis. 
  A thin black line extends upwards from the rod COM, and the COM of the particulate is shown by a white and black disk.   
  More particulate is heaped on the right side of the rod for positive $\theta$ and more on the left side for negative $\theta$. 
  This uneven particulate generates a restoring torque. 
 We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$.

\subsection{Force and Torque}
%
By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
\begin{figure}
\centering
\renewcommand{\figwid}{0.8\columnwidth}
\begin{overpic}[width =\figwid]{AngleOfRepose.pdf}
\put(39.1,12){\scriptsize $\leftarrow \, \ell \, \, \rightarrow$}
\rotatebox{22.5}{\put(55,-12){\scriptsize $\leftarrow \, \ell \, \, \rightarrow$}}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:AngleOfReposeForce} Top plot shows colored particulate heaped up on pink-colored long rods. 
 Middle plot shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
   The maximum torque values from \eqref{eq:maxTorqueAngleGivenAOR} are shown with black dots, producing a line that is approximately $-\ell^3/36 \theta_{t_{\max}}$.
% Generating code is in the attachment. 
\vspace{-2em}
}
\end{figure}
%
\begin{align}
F(\theta,\alpha,\ell) &=  \!\!\!\!\!\!\!\! \int\limits_{-\cos{\theta}\ell/2}\limits^{\cos{\theta}\ell/2} \!\!\!\!\!\!\!\! x\left(
 \min\!\! \left(\!\! 
    \frac{ \ell\sin{\alpha \pm\theta }}{\cos{\alpha}}  \mp x \tan{\alpha} \!
    %, \frac{ \ell\sin{\alpha-\theta }}{\cos{\alpha}}  + x \tan{\alpha}
 \right)
 \!-\! x \tan{\theta}\!
 \right)  dx
 \nonumber \\
F(\theta,\alpha,\ell) &=\left\{
\begin{array}{ll}
\frac{\ell^2\Big(\cos(2\theta)-\cos(2\alpha)\Big)}{8\cos(\alpha)\sin(\theta)} &   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise .}
\end{array} 
\right . \label{eq:ForceAOR}
\end{align}

@@

The force for different angle of repose values are shown in the middle plot of Fig.~\ref{fig:AngleOfReposeForce}, with the rod length $\ell=1$. 
Torque will also be similarly defined as
\begin{align}
\tau(\theta, \alpha, \ell) =\left\{
\begin{array}{ll}
\frac{\ell^3\Big(\cos{2\alpha}-\cos{2\theta}\Big)\sin{\theta} }{48\sin^2(\alpha)}&   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise .} 
\end{array} 
\right.  \label{eq:TorqueAOR}
\end{align}
Torque is shown in the bottom plot of Fig.~\ref{fig:AngleOfReposeForce} with the rod length $\ell=1$. 
 Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.
Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$ and is defined as
\begin{align}\label{eq:maxTorqueAngleGivenAOR}
\theta_{t_{\max}} = \frac{\sin(\alpha)}{\sqrt{3}}.
\end{align}
To maximize the torque a particulate swarm applies on a thin rod, the swarm should move in the direction $-\theta_{t_{\max}} - 90^\circ$ with respect to the long axis of the rod.

>>>>

Consider a swarm of granular particles applying force to a rod. 
If the rod moves slower than the particles, some particles will slide past the rod, but other particles will build up behind the rod in a characteristic triangular shape %defined by an apex angle. 
 defined by a steepest angle of descent perpendicular to the direction of particle motion. 
This piling up is common to all granular media, and the angle formed perpendicular to the angle of attack $\alpha$ is the \emph{angle of repose}\footnote{angle of repose is a 3D phenomenon, but this paper only considers the 2D projection of the piled particles.}.
$\alpha$ is a function  of many variables, including the physical shape and composition of the particles and the external force applied.  
 Three different values of angle of repose are shown in Fig.~\ref{fig:angle}. The center of mass (COM) of the rod is in the middle of the rod, but  the COM of the granular particles changes for different $\alpha$ values. % for the media, see \cite{angleofRepose}. 
 The $\alpha$ values can be measured experimentally.   Four types of particles with different angles of repose are shown in Fig.~\ref{fig:AngleOfReposeForce}(a). 
  Once the angle of repose is known, we can estimate the force and torque that the particles are applying to the rod as a function of the rod's length, the angle of repose, and orientation of the rod.
 In this plot, particulate is moving in the $-y$ direction, and the rods are tilted at $\theta=\{-45,-22.5,0,22.5,45\}^\circ$ with respect to the $x$-axis. 
  A thin black line extends upwards from the rod COM, and the COM of the particulate is shown by a white and black disk.   
  More particulate is heaped on the right side of the rod for positive $\theta$ and more on the left side for negative $\theta$. 
  This uneven particulate generates a restoring torque. 
 We define the angle of repose as $\alpha$, the rod's orientation relative to 90$^\circ$ from the particle movement vector as $\theta$, and the rod's length as $\ell$.

\subsection{Force and Torque}
%
By integrating over the triangular shape, the force applied to the rod (when a unit area of particles produces 1 N of force) is
\begin{figure}
\centering
\renewcommand{\figwid}{0.8\columnwidth}
\begin{overpic}[width =\figwid]{AngleOfRepose.pdf}
\put(0,55){(a)}
\put(39.1,12){\scriptsize $\leftarrow \, \ell \, \, \rightarrow$}
\rotatebox{22.5}{\put(55,-12){\scriptsize $\leftarrow \, \ell \, \, \rightarrow$}}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =\figwid]{AngleOfReposeForce.pdf}%\put(1,55){a)}
\put(0,45){(b)}
\end{overpic}\\
\vspace{0.5em}
\begin{overpic}[width =\figwid]{AngleOfReposeTorque.pdf}%\put(1,55){a)}
\put(0,55){(c)}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:AngleOfReposeForce} (a) plots colored particulate heaped up on pink-colored long rods. 
 (b) shows the force applied to the rod and bottom the torque as a function of $\theta$ for four angle of repose values.
  (c)  The maximum torque values from \eqref{eq:maxTorqueAngleGivenAOR} are shown with black dots, producing a line that is approximately $-\ell^3/36 \theta_{t_{\max}}$.
% Generating code is in the attachment. 
\vspace{-2em}
}
\end{figure}
%
\begin{align}
F(\theta,\alpha,\ell) &=  \!\!\!\!\!\!\!\! \int\limits_{-\cos{\theta}\ell/2}\limits^{\cos{\theta}\ell/2} \!\!\!\!\!\!\!\! x\left(
 \min\!\! \left(\!\! 
    \frac{ \ell\sin{\alpha \pm\theta }}{\cos{\alpha}}  \mp x \tan{\alpha} \!
    %, \frac{ \ell\sin{\alpha-\theta }}{\cos{\alpha}}  + x \tan{\alpha}
 \right)
 \!-\! x \tan{\theta}\!
 \right)  dx
 \nonumber \\
F(\theta,\alpha,\ell) &=\left\{
\begin{array}{ll}
\frac{\ell^2\Big(\cos(2\theta)-\cos(2\alpha)\Big)}{8\cos(\alpha)\sin(\theta)} &   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise .}
\end{array} 
\right . \label{eq:ForceAOR}
\end{align}

@@

The force for different angle of repose values are shown in Fig.~\ref{fig:AngleOfReposeForce}(a), with the rod length $\ell=1$. 
Torque will also be similarly defined as
\begin{align}
\tau(\theta, \alpha, \ell) =\left\{
\begin{array}{ll}
\frac{\ell^3\Big(\cos{2\alpha}-\cos{2\theta}\Big)\sin{\theta} }{48\sin^2(\alpha)}&   -\alpha<\theta<\alpha\\
0 &    \textrm{otherwise .} 
\end{array} 
\right.  \label{eq:TorqueAOR}
\end{align}
Torque is shown in Fig.~\ref{fig:AngleOfReposeForce}(c) with the rod length $\ell=1$. 
 Given sufficient particles to pile up to the angle of repose, this torque tends to stabilize the object to be perpendicular to the pushing direction.
Force is maximized with $\theta=0$, but the $\theta$ value that maximizes torque is a function of $\alpha$ and is defined as
\begin{align}\label{eq:maxTorqueAngleGivenAOR}
\theta_{t_{\max}} = \frac{\sin(\alpha)}{\sqrt{3}}.
\end{align}
To maximize the torque a particulate swarm applies on a thin rod, the swarm should move in the direction $-\theta_{t_{\max}} - 90^\circ$ with respect to the long axis of the rod.
